---
title: "Appendix A: Results of Structural Equation Models (SEMs)"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib  
---

\def\figurename{Figure A}
\def\tablename{Table A}

```{r setup, include=FALSE}
# Libraries
library(tidyverse)
library(brms)
library(dagitty)
library(broom.mixed)

# Knitr Settings
knitr::opts_chunk$set(echo = FALSE, fig.width = 9, fig.height = 5, warning = FALSE)

# Help function to prepare the results tables
getres <- function(fit) {
  res <- fit %>% 
    tidy %>% 
    filter(effect == "fixed") %>% 
    filter(term != "(Intercept)")
  res$Equation <- nrformula[match(res$response, nameformula)]
  res$response <- nameVariable[match(res$response, abbrVariable)]
  res$term <- nameVariable[match(res$term, abbrVariable)]
  res %>% 
    transmute(
      Equation, 
      Response = response,
      Predictor = term,
      Coefficient = estimate,
      "2.5%-Quantile" = conf.low,
      "97.5%-Quantile" = conf.high
    )
}
```

```{r dataselection}

# Load haymeadow data
dat <- read_csv("Data-Raw/Data_sites.csv", show_col_types = FALSE)
load("Data-Raw/community_matrix.RData")

# Data selection
ausw <- 
  (dat$Delarze == "4.5.1" | dat$Delarze == "4.5.2") &
  !is.na(dat$Delarze) &
  !is.na(dat$Ellenberg_R) #&
  # !(dat$Programm == "BDM" & dat$SR < 35) &
  # !(dat$Programm == "ALL-EMA" & is.na(match(dat$BFF, c(611, 851, 634, 852, 921611, 922611, 924611, 921852))))
dat <- dat[ausw,] 
comdat <- comdat[ausw, ]

# Transfrom vairables
dat <- dat %>% 
  mutate(
    EUNIS = ifelse(elevation < 600, "R22", "R23"),
    ndep = (nitrogen_depostion - 10) / 20,
    ele = (elevation - 1000) / 1000,
    temp = (temperature - 10) / 10,
    preci = (precipitation - 1000) / 1000,
    light = (Ellenberg_L - 3.5),
    pH = (Ellenberg_R - 3.3),
    soilN = Ellenberg_N - 3,
    NDVI = (NDVI - 6000) / 1000,
    LUI = landuse_intensity-0.5,
    specpool = (species_pool - 250) / 10,
    incli = (inclination - 10) / 5,
    BDM = as.integer(Programm == "BDM")
  )
```

# Introduction
In this appendix the detailed SEM results are provided from applying differnt models to the data of the low and medium-altitude hay meadows (EUNIS R2.2) and the mountain hay meadows (R2.3). As measure of species richness either the total species richness or the number of the target species were used.

# Main model
In this chapter we applied the data to the conceptual model as described in the section *conceptual model* in the main body of the manuscript for a detailed description of the model. The main model is depicted as a causal diagram in the following figure. 

```{r conceptualmodel, fig.cap = "Conceptual model used in the SEM."}

# Model description according to http://www.dagitty.net/
conceptualmod <- dagitty('
dag {
bb="-5.227,-3.153,4.663,3.021"
"Agricultural landuse intensity" [pos="-2.670,-1.256"]
"Nitrogen deposition" [exposure,pos="-4.317,-2.070"]
"Plant species richness" [outcome,pos="0.747,-1.471"]
"Soil pH" [pos="0.709,-2.095"]
"Soil nutrient content" [pos="-2.795,-1.763"]
"Species pool" [pos="-1.211,-1.085"]
Light [pos="-1.411,-1.427"]
NDVI [pos="-0.974,-1.739"]
Precipitation [pos="-4.229,-1.436"]
Temperature [pos="-4.242,-1.119"]
"Agricultural landuse intensity" -> "Nitrogen deposition"
"Agricultural landuse intensity" -> "Plant species richness"
"Agricultural landuse intensity" -> "Soil nutrient content"
"Agricultural landuse intensity" -> Light
"Nitrogen deposition" -> "Plant species richness" [pos="0.173,-2.085"]
"Nitrogen deposition" -> "Soil pH"
"Nitrogen deposition" -> "Soil nutrient content"
"Soil pH" -> "Plant species richness"
"Soil nutrient content" -> "Plant species richness" [pos="-0.538,-2.041"]
"Soil nutrient content" -> NDVI
"Species pool" -> "Plant species richness"
Light -> "Plant species richness"
NDVI -> "Plant species richness"
NDVI -> Light
Precipitation -> "Agricultural landuse intensity"
Precipitation -> "Nitrogen deposition"
Precipitation -> NDVI
Temperature -> "Agricultural landuse intensity"
Temperature -> "Species pool"
Temperature -> NDVI
}
')

# Plot conceptual model
plot(conceptualmod)
```

\newpage
Based on the conceptual model the following equations were used to analyse the data:

\begin{align}
  Plant\ species\ richness &= \begin{aligned}[t] &Nitrogen\ deposition + Soil\ alkalinity + Soil\ nutrient\ content + NDVI + \\ &Light + Landuse\ intensity + Species\ pool \end{aligned} \\
  Soil\ alkalinity &= Nitrogen\ deposition \\
  Soil\ nutrient\ content &= Nitrogen\ deposition + Landuse\ intensity \\
  NDVI &= Soil\ nutrient\ content + Precipitation + Temperature \\
  Light &= NDVI + Landuse\ intensity \\
  Nitrogen\ deposition &= Precipitation + Landuse\ intensity \\
  Landuse\ intensity &= Temperature + Precipitation \\
  Species\ pool &= Temperature 
\end{align}

```{r modformulas}
# brms formulas
mod_sr <- bf(SR ~ ndep + pH + soilN + NDVI + light + LUI + specpool, family = poisson)
mod_ph <- bf(pH ~ ndep)
mod_soilN <- bf(soilN ~ ndep + LUI)
mod_ndvi <- bf(NDVI ~ soilN + preci + temp)
mod_light <- bf(light ~ NDVI + LUI)
mod_ndep <- bf(ndep ~ preci + LUI)
mod_LUI <- bf(LUI ~ temp + preci)
mod_spec_pool <- bf(specpool ~ temp)

# Formulas
nrformula <- c("(1)", "(2)",  "(3)", "(4)", "(5)", "(6)", "(7)", "(8)")
nameformula <- c("SR", "pH", "soilN", "NDVI", "light", "ndep", "LUI", "specpool")

# Variable names
nameVariable = c("Plant species richness", "Nitrogen deposition", "Soil alkalinity", "Soil nutrient content", "NDVI", "Light", "Precipitation", "Temperature", "Landuse intensity", "Species pool")
abbrVariable = c("SR", "ndep", "pH", "soilN", "NDVI", "light", "preci", "temp", "LUI", "specpool")
```

These formulas were implemented and applied to the data using the R-package `brms` [@brms]. The results are shown in the following tables. The same results are also depicted in Figure 3 in the main body of the manuscript.
\newpage
```{r r22sr}
# R22, total species richness
r22sr_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(SR)),
  file = "Modres/main_model/SEM-R22-SR"
) 

# Res Table
getres(r22sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **total species richness** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```
\newpage
```{r r22uzl}
# R22, target species richness
r22uzl_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/main_model/SEM-R22-UZL"
) 

# Res Table
getres(r22uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **numer of target species** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```
\newpage
```{r r23sr}
# R23, total species richness
r23sr_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(SR)),
  file = "Modres/main_model/SEM-R23-SR"
) 

# Res Table
getres(r23sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **total species richness** applied to the to the data of the **lmountain altitude hay meadows** (EUNIS R2.3).")
```
\newpage
```{r r23uzl}
# R23, target species richness
r23uzl_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/main_model/SEM-R23-UZL"
) 

# Res Table
getres(r23uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **numer of target species** applied to the to the data of the **mountain altitude hay meadows** (EUNIS R2.3).")
```

\newpage
# Starting model
We started the analyses with a simpler model that is shown in the following graph. Only later we adapted this starting model based on the graphical model by @demalach2018toward to the main model. The results of this simpler model probably has shaped our conclusion. Accordingly, we also present the results of this model here. This simpler model is a selection of the variables from the main model, except for Inclination. In this simpler model we used inclination of the terrain at the sampling site as a proxi for the landuse intensity (higher inclination = lower landuse intensity). In the main model we then replaced inclination by the landuse intensity index. 

```{r, fig.cap = "Conceptual model used in the SEM."}

# Model description according to http://www.dagitty.net/
conceptualmod <- dagitty('
dag {
bb="-5.227,-3.153,4.663,3.021"
"Nitrogen deposition" [exposure,pos="-4.303,-2.504"]
"Plant species richness" [outcome,pos="-0.720,-2.261"]
"Soil alkalinity" [pos="-2.147,-2.641"]
Inclination [pos="-4.465,-1.514"]
Light [pos="-0.987,-1.753"]
NDVI [pos="-2.292,-1.544"]
Precipitation [pos="-4.497,-1.812"]
Temperature [pos="-4.506,-2.100"]
"Nitrogen deposition" -> "Plant species richness"
"Nitrogen deposition" -> "Soil alkalinity"
"Nitrogen deposition" -> NDVI
"Soil alkalinity" -> "Plant species richness"
Inclination -> "Plant species richness"
Inclination -> NDVI
Light -> "Plant species richness"
NDVI -> "Plant species richness"
NDVI -> Light
Precipitation -> "Plant species richness"
Precipitation -> NDVI
Temperature -> "Plant species richness"
Temperature -> NDVI
}
')

# Plot conceptual model
plot(conceptualmod)
```

\newpage
Based on the conceptual model the following equations were used to analyse the data:

\begin{align}
  Plant\ species\ richness &= \begin{aligned}[t] &Nitrogen\ deposition + Soil\ alkalinity + Temperature + Precipitation + \\ &Inclination + NDVI + Light \end{aligned} \\
  Soil\ alkalinity &= Nitrogen\ deposition \\
  NDVI &= Nirogen\ deposition + Temperature + Precipitation + Inclination\\
  Light &= NDVI
\end{align}

```{r}
# brms formulas
mod_sr <- bf(SR ~ ndep + pH + temp + preci + incli + NDVI + light, family = poisson)
mod_ph <- bf(pH ~ ndep)
mod_ndvi <- bf(NDVI ~ ndep + temp + preci + incli)
mod_light <- bf(light ~ NDVI)

# Formulas
nrformula <- c("(17)", "(18)",  "(19)", "(20)", "(21)")
nameformula <- c("SR", "pH", "NDVI", "light")

# Variable names
nameVariable = c("Plant species richness", "Nitrogen deposition", "Soil alkalinity", "NDVI", "Light", "Precipitation", "Temperature", "Inclination")
abbrVariable = c("SR", "ndep", "pH", "NDVI", "light", "preci", "temp", "incli")
```

These formulas were implemented and applied to the data using the R-package `brms` [@brms]. The results are shown in the following tables. 

```{r}
# R22, total species richness
r22sr_SEM <- brm(
  mod_sr  + mod_ph + mod_ndvi + mod_light + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(SR)),
  file = "Modres/firstmod/SEM-R22-SR"
) 

# Res Table
getres(r22sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model depicted in Fig. 3 for the **total species richness** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```
\newpage
```{r}
# R22, target species richness
r22uzl_SEM <- brm(
  mod_sr  + mod_ph + mod_ndvi + mod_light + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/firstmod/SEM-R22-UZL"
) 

# Res Table
getres(r22uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model depicted in Fig. 3 for the **numer of target species** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```
\newpage
```{r}
# R23, total species richness
r23sr_SEM <- brm(
  mod_sr  + mod_ph + mod_ndvi + mod_light + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(SR)),
  file = "Modres/firstmod/SEM-R23-SR"
) 

# Res Table
getres(r23sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model depicted in Fig. 3 for the **total species richness** applied to the to the data of the **lmountain altitude hay meadows** (EUNIS R2.3).")
```
\newpage
```{r}
# R23, target species richness
r23uzl_SEM <- brm(
  mod_sr  + mod_ph + mod_ndvi + mod_light + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/firstmod/SEM-R23-UZL"
) 

# Res Table
getres(r23uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model depicted in Fig. 3 for the **numer of target species** applied to the to the data of the **mountain altitude hay meadows** (EUNIS R2.3).")
```


# Correlation of predictor variables

The following two figures show the pairwise correlation matrix of the predictor variables that we used in the main model. 

```{r, fig.cap = "Pairwise correlation plot using the data of the lowland hay meadows."}
## put histograms on the diagonal
panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

## put (absolute) correlations on the upper panels,
## with size proportional to the correlations.
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y, use = "pairwise.complete.obs"))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}

dat %>% 
  filter(EUNIS == "R22") %>% 
  transmute(ndep, preci, temp, soilN, LUI, NDVI, light, pH, specpool) %>% 
  pairs(lower.panel = panel.smooth, upper.panel = panel.cor, gap=0, row1attop = FALSE, pch = 16, cex = 0.2, cex.cor = 3)
```

```{r, fig.cap = "Pairwise correlation plot using the data of the mountain hay meadows."}
dat %>% 
  filter(EUNIS == "R23") %>% 
  transmute(ndep, preci, temp, soilN, LUI, NDVI, light, pH, specpool) %>% 
  pairs(lower.panel = panel.smooth, upper.panel = panel.cor, gap=0, row1attop = FALSE, pch = 16, cex = 0.2, cex.cor = 3)
```

# References
