---
title: "Appendix A: SEM Results of main model"
output: pdf_document 
abstract: "Appendix A for the MS The Sensitivity of hay meadow species diversity to nitrogen deposition"
bibliography: references.bib  
---

```{r setup, include=FALSE}
# Libraries
library(tidyverse)
library(brms)
library(dagitty)
library(broom.mixed)

# Knitr Settings
knitr::opts_chunk$set(echo = FALSE, fig.width = 9, fig.height = 5, warning = FALSE)

# Help function to prepare the results tables
getres <- function(fit) {
  res <- fit %>% 
    tidy %>% 
    filter(effect == "fixed") %>% 
    filter(term != "(Intercept)")
  res$Equation <- nrformula[match(res$response, nameformula)]
  res$response <- nameVariable[match(res$response, abbrVariable)]
  res$term <- nameVariable[match(res$term, abbrVariable)]
  res %>% 
    transmute(
      Equation, 
      Response = response,
      Predictor = term,
      Coefficient = estimate,
      "2.5%-Quantile" = conf.low,
      "97.5%-Quantile" = conf.high
    )
}
```

```{r dataselection}

# Load haymeadow data
dat <- read_csv("Data-Raw/Data_sites.csv", show_col_types = FALSE)
load("Data-Raw/community_matrix.RData")

# Data selection
ausw <- 
  (dat$Delarze == "4.5.1" | dat$Delarze == "4.5.2") &
  !is.na(dat$Delarze) &
  !is.na(dat$Ellenberg_R) &
  !(dat$Programm == "BDM" & dat$SR < 35) &
  !(dat$Programm == "ALL-EMA" & is.na(match(dat$BFF, c(611, 851, 634, 852, 921611, 922611, 924611, 921852))))
dat <- dat[ausw,] 
comdat <- comdat[ausw, ]

# Transfrom vairables
dat <- dat %>% 
  mutate(
    EUNIS = ifelse(elevation < 600, "R22", "R23"),
    ndep = (nitrogen_depostion - 10) / 20,
    ele = (elevation - 1000) / 1000,
    temp = (temperature - 10) / 10,
    preci = (precipitation - 1000) / 1000,
    light = (Ellenberg_L - 3.5),
    pH = (Ellenberg_R - 3.3),
    soilN = Ellenberg_N - 3,
    NDVI = (NDVI - 6000) / 1000,
    LUI = landuse_intensity-0.5,
    specpool = (species_pool - 250) / 10
  )
```

# Introduction
In this appendix the detailed SEM results are provided from applying the main model to the data of the low and medium-altitude hay meadows (EUNIS R2.2) and the data of the mountain hay meadows (R2.3). As measure of species richness either the total species richness or the number of the target species were used.
See the section *conceptual model* in the main body of the manuscript for a detailed description of the model. The model is depicted as a causal diagram in the following figure. 

```{r conceptualmodel, fig.cap = "Conceptual model used in the SEM."}

# Model description according to http://www.dagitty.net/
conceptualmod <- dagitty('
dag {
bb="-5.227,-3.153,4.663,3.021"
"Landuse intensity" [pos="-2.670,-1.256"]
"Nitrogen deposition" [exposure,pos="-4.317,-2.070"]
"Plant species richness" [outcome,pos="0.747,-1.471"]
"Soil alkalinity" [pos="0.709,-2.095"]
"Soil nutrient content" [pos="-2.795,-1.763"]
"Species pool" [pos="-1.211,-1.085"]
Light [pos="-1.411,-1.427"]
NDVI [pos="-0.974,-1.739"]
Precipitation [pos="-4.229,-1.436"]
Temperature [pos="-4.242,-1.119"]
"Landuse intensity" -> "Nitrogen deposition"
"Landuse intensity" -> "Plant species richness"
"Landuse intensity" -> "Soil nutrient content"
"Landuse intensity" -> Light
"Nitrogen deposition" -> "Plant species richness" [pos="0.173,-2.085"]
"Nitrogen deposition" -> "Soil alkalinity"
"Nitrogen deposition" -> "Soil nutrient content"
"Soil alkalinity" -> "Plant species richness"
"Soil nutrient content" -> "Plant species richness" [pos="-0.538,-2.041"]
"Soil nutrient content" -> NDVI
"Species pool" -> "Plant species richness"
Light -> "Plant species richness"
NDVI -> "Plant species richness"
NDVI -> Light
Precipitation -> "Landuse intensity"
Precipitation -> "Nitrogen deposition"
Precipitation -> NDVI
Temperature -> "Landuse intensity"
Temperature -> "Species pool"
Temperature -> NDVI
}
')

# Plot conceptual model
plot(conceptualmod)

```

\newpage
Based on the conceptual model the following equations were used to analyse the data:

\begin{align}
  Plant\ species\ richness &= \begin{aligned}[t] &Nitrogen\ deposition + Soil\ alkalinity + Soil\ nutrient\ content + NDVI + \\ &Light + Landuse\ intensity + Species\ pool \end{aligned} \\
  Soil\ alkalinity &= Nitrogen\ deposition \\
  Soil\ nutrient\ content &= Nitrogen\ deposition + Landuse\ intensity \\
  NDVI &= Soil\ nutrient\ content + Precipitation + Temperature \\
  Light &= NDVI + Landuse\ intensity \\
  Nitrogen\ deposition &= Precipitation + Landuse\ intensity \\
  Landuse\ intensity &= Temperature + Precipitation \\
  Species\ pool &= Temperature 
\end{align}

# Model Results

```{r modformulas}
# brms formulas
mod_sr <- bf(SR ~ ndep + pH + soilN + NDVI + light + LUI + specpool, family = poisson)
mod_ph <- bf(pH ~ ndep)
mod_soilN <- bf(soilN ~ ndep + LUI)
mod_ndvi <- bf(NDVI ~ soilN + preci + temp)
mod_light <- bf(light ~ NDVI + LUI)
mod_ndep <- bf(ndep ~ preci + LUI)
mod_LUI <- bf(LUI ~ temp + preci)
mod_spec_pool <- bf(specpool ~ temp)

# Formulas
nrformula <- c("(1)", "(2)",  "(3)", "(4)", "(5)", "(6)", "(7)", "(8)")
nameformula <- c("SR", "pH", "soilN", "NDVI", "light", "ndep", "LUI", "specpool")

# Variable names
nameVariable = c("Plant species richness", "Nitrogen deposition", "Soil alkalinity", "Soil nutrient content", "NDVI", "Light", "Precipitation", "Temperature", "Landuse intensity", "Species pool")
abbrVariable = c("SR", "ndep", "pH", "soilN", "NDVI", "light", "preci", "temp", "LUI", "specpool")
```

The formulas were applied using the R-package `brms` [@brms]. The results are shown in the following tables. 

```{r r22sr}
# R22, total species richness
r22sr_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(SR)),
  file = "Modres/main_model/SEM-R22-SR"
) 

# Res Table
getres(r22sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **total species richness** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```

```{r r22uzl}
# R22, target species richness
r22uzl_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R22") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/main_model/SEM-R22-UZL"
) 

# Res Table
getres(r22uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **numer of target species** applied to the to the data of the **low and medium-altitude hay meadows** (EUNIS R2.2).")
```
\newpage
```{r r23sr}
# R23, total species richness
r23sr_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(SR)),
  file = "Modres/main_model/SEM-R23-SR"
) 

# Res Table
getres(r23sr_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **total species richness** applied to the to the data of the **lmountain altitude hay meadows** (EUNIS R2.3).")
```
\newpage
```{r r23uzl}
# R23, target species richness
r23uzl_SEM <- brm(
  mod_sr  + mod_ndvi + mod_light + mod_ph + mod_soilN + mod_ndep + mod_LUI + mod_spec_pool + set_rescor(FALSE),
  data = dat %>% filter(EUNIS == "R23") %>% mutate(SR = as.integer(UZL)),
  file = "Modres/main_model/SEM-R23-UZL"
) 

# Res Table
getres(r23uzl_SEM) %>% 
  knitr::kable(
    digits = 2, 
    caption = "SEM Results from applyig the structural equation model for the **numer of target species** applied to the to the data of the **mountain altitude hay meadows** (EUNIS R2.3).")
```



# References
